---
title: "PSTAT 175 Final Project"
author: "Xiaoxi Guo, Zhuoqi Niu, Kelly Wang, David Zhang"
date: "12/5/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Dataset
```{r}
NKI <- read.csv("NKI_cleaned.csv")
NKI <- NKI[,c('Patient','age','eventdeath','survival','timerecurrence','chemo','hormonal','amputation','histtype','diam','posnodes','grade','angioinv')]
head(NKI)
```

#Load Packages
```{r}
library(survival)
library(survminer)
```

#Data Exploration
```{r}
summary(NKI[,c('age','survival','timerecurrence')])
lapply(NKI[,c('chemo','hormonal','amputation','grade')], function(x) {
  return(table(x))
})
```

```{r}
quantile(NKI$survival)
plot(density(NKI$survival), main='Survival Time Distribution', xlab='Years')
polygon(density(NKI$survival), col='darkseagreen2', border='darkseagreen4')
```

```{r}
quantile(NKI$age)
plot(density(NKI$age), main='Age Distribution', xlab='Age')
polygon(density(NKI$age), col='darkseagreen2', border='darkseagreen4')
```

##Group by approximately 25% and 75% quantile
```{r}
NKI$diamgroup[NKI$diam<=15] = 1
NKI$diamgroup[NKI$diam>15 & NKI$diam<30] = 2
NKI$diamgroup[NKI$diam>=30] = 3

NKI$agegroup[NKI$age<=40.75] = 1
NKI$agegroup[NKI$age>40.75 & NKI$age<49.00] = 2
NKI$agegroup[NKI$age>=49.00] = 3
```

```{r}
head(NKI) #Check first rows of dataset
```
Covariates that we are interested are *chemo*, *hormonal*, *amputation*, *diamgroup*, *agegroup*, *grade*.\newline

#KM Plots
```{r}
NKI.surv <- Surv(NKI$survival,NKI$eventdeath)
NKI.fit <- survfit(NKI.surv ~ 1)
plot(NKI.fit, main = "Kaplan-Meier Curve (General)", xlab = "Percentage", ylab = "Time After Diagnosis (Years)")
```

```{r}
km_chemo = survfit(NKI.surv~chemo,data=NKI)
par(mar=c(5,5,4,2))
plot(km_chemo, xlab="Survival time",ylab = expression(hat(S)(t)),main = "KM Curve (Chemo and No Chemo)",lwd=2, col=1:2, mark.time = TRUE,mark=18)
legend("topright",legend=c("no chemo","chemo"),col=1:2,pch=rep(19,2))
```

```{r km curve hormonal}
km_hormonal = survfit(NKI.surv~hormonal,data=NKI)
par(mar=c(5,5,4,2))
plot(km_hormonal, xlab="Survival time",ylab = expression(hat(S)(t)),main = "KM Curve (hormonal)",lwd=2, col=1:2, mark.time = TRUE,mark=18)
legend("bottomleft",legend=c("no hormonal","hormonal"),col=1:2,pch=rep(19,2))
```

```{r km curve amputation}
km_amputation = survfit(NKI.surv~amputation,data=NKI)
par(mar=c(5,5,4,2))
plot(km_amputation, xlab="Survival time",ylab = expression(hat(S)(t)),main = "KM Curve (Amputation)",lwd=2, col=1:2, mark.time = TRUE,mark=18)
legend("topright",legend=c("no amputation","amputation"),col=1:2,pch=rep(19,2))
```

```{r grade level}
km_grade = survfit(NKI.surv~grade,data=NKI)
par(mar=c(5,5,4,2))
plot(km_grade, xlab="Survival time",ylab = expression(hat(S)(t)),main = "KM Curve (grade)",lwd=3, col=1:3, mark.time = TRUE,mark=18)
legend("topright",legend=c("1","2","3"),col=1:3,pch=rep(19,2))
```

```{r diamgroup level}
km_diamgroup = survfit(NKI.surv~diamgroup,data=NKI)
par(mar=c(5,5,4,2))
plot(km_diamgroup, xlab="Survival time",ylab = expression(hat(S)(t)),main = "KM Curve (diamgroup)",lwd=3, col=1:3, mark.time = TRUE,mark=18)
legend("topright",legend=c("1","2","3"),col=1:3,pch=rep(19,2))
```

```{r agegroup level}
km_agegroup = survfit(NKI.surv~agegroup,data=NKI)
par(mar=c(5,5,4,2))
plot(km_agegroup, xlab="Survival time",ylab = expression(hat(S)(t)),main = "KM Curve (agegroup)",lwd=3, col=1:3, mark.time = TRUE,mark=18)
legend("topright",legend=c("1","2","3"),col=1:3,pch=rep(19,2))
```

#LogRank Test
```{r}
survdiff(NKI.surv~NKI$chemo) 
survdiff(NKI.surv~NKI$hormonal) 
survdiff(NKI.surv~NKI$amputation) 
survdiff(NKI.surv~NKI$grade) 
survdiff(NKI.surv~NKI$agegroup)
survdiff(NKI.surv~NKI$diamgroup)
```

#Model Selection
##Model1: Backward selection
```{r}
cox <- coxph(Surv(NKI$survival, NKI$eventdeath)~diamgroup+grade+agegroup, data = NKI)
step(cox, direction = "backward")
```

##Model2: Likelihood tests selection
```{r cox zph}
cox1 <- coxph(Surv(NKI$survival, NKI$eventdeath)~diamgroup+grade+agegroup, data = NKI)
anova(cox1)
```
All variables are significant (using 0.1 as critical value).

##Compared model
```{r}
fit0 <- coxph(Surv(survival, eventdeath)~grade, data = NKI)
fit1 <- coxph(Surv(survival, eventdeath)~grade+diamgroup, data = NKI)
lrt1 = 2*(fit1$loglik[2]-fit0$loglik[2])
pchisq(lrt1,df=1,lower.tail = FALSE)

fit2 <- coxph(Surv(survival, eventdeath)~grade+diamgroup+agegroup, data = NKI)
lrt2 = 2*(fit2$loglik[2]-fit1$loglik[2])
pchisq(lrt2,df=1,lower.tail = FALSE)
```
Do not mention the specific numbers from the compared model part; Give our final decision only since the numbers do not look beautiful.\newline

#Model Checking
##Method1: Residual tests
```{r}
NKI <- within(NKI, {
  grade <- factor(grade, labels = c("1", "2","3"))
  diamgroup <- factor(diamgroup, labels = c("1", "2","3"))
  agegroup <- factor(agegroup, labels = c("1", "2","3"))
})

cox1 <- coxph(Surv(NKI$survival, NKI$eventdeath)~diamgroup+grade+agegroup, data = NKI)

cox.zph(cox1)
```
Since p-value for *grade* is less than 0.05, we need to stratify it.

```{r cox zphn stratify}
cox2 <- coxph(Surv(NKI$survival, NKI$eventdeath)~diamgroup+strata(grade)+agegroup, data = NKI)
cox.zph(cox2)
```

##Method2:C-log-log Plot
```{r c log log}
#grade
plot(survfit(NKI.surv ~ NKI$grade),lwd=2,col=1:3, fun="cloglog",main = "cloglog grade",xlab="Time until Death \n in years",ylab="log(S(t))") 
legend('topleft',c("1","2","3"),fill = 1:3)
#diamgroup
plot(survfit(NKI.surv ~ NKI$diamgroup),lwd=2,col=1:3, fun="cloglog",main = "cloglog diamgroup",xlab="Time until Death \n in years",ylab="log(S(t))") 
legend('topleft',c("1","2","3"),fill = 1:3)
#agegroup
plot(survfit(NKI.surv ~ NKI$agegroup),lwd=2,col=1:3, fun="cloglog",main = "cloglog agegroup",xlab="Time until Death \n in years",ylab="log(S(t))") 
legend('topleft',c("1","2","3"),fill = 1:3)
```

#Interaction Terms
```{r}
coxA <- coxph(Surv(NKI$survival, NKI$eventdeath)~diamgroup*strata(agegroup), data = NKI)
coxB <- coxph(Surv(NKI$survival, NKI$eventdeath)~diamgroup*strata(grade), data = NKI)
coxC <- coxph(Surv(NKI$survival, NKI$eventdeath)~strata(agegroup)*strata(grade), data = NKI)
anova(coxA)
anova(coxB)
anova(coxC)
```
All the interaction terms have large p-values (greater than 0.05). Therefore, we will not include interaction terms in our model.\newline

#Hazard Ratios and C.I.
```{r}
cox_ns <- coxph(Surv(NKI$survival, NKI$eventdeath)~diamgroup, data = NKI)
ggforest(cox_ns, data = NULL, main = "Hazard ratio",
         cpositions = c(0.02, 0.22, 0.4), fontsize = 0.7,
         refLabel = "reference", noDigits = 2)
```

#Baseline Hazard Rates
```{r}
#Baseline Hazard Plot for grade
fit_grade <- coxph(Surv(survival, eventdeath)~strata(grade), data=NKI)
by_grade <- basehaz(fit_grade) %>% 
  group_by(strata)
ggplot(by_grade, aes(x = time, y = hazard)) + 
  geom_line(aes(color = strata), size=2) +
  ggtitle("Baseline Hazard Rates for covariate grade")

#Baseline Hazard Plot for agegroup
fit_age <- coxph(Surv(survival, eventdeath)~strata(agegroup), data=NKI)
by_age <- basehaz(fit_age) %>% 
  group_by(strata)
ggplot(by_age, aes(x = time, y = hazard)) + 
  geom_line(aes(color = strata), size=2) +
  ggtitle("Baseline Hazard Rates for covariate agegroup")

```













